---
title: "Final Modeling Notebook Sahel"
author: "Ignacio García Sánchez-Migallón"
date: "20/8/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('SASmixed')
library(nlme)
library(lme4)
library(ggplot2)
library(grid)
library(gridExtra)
library(gstat)
library(tidyverse)
library(fasterize)
library(tiff)
library(raster)
library(sf)
library(spatialreg)
library(spdep)
rm(list=ls())

```

## Introduction

Perform an exploratory analysis of the data.
All the series of plots generated with grid. have to be done in the console as to obtain bigger plots.

```{r}


setwd("C:/Users/IgnacioGarcia/Desktop/UNIVERSIDAD/Statistics for data science/Master thesis/data")

data = st_read('Mixed_model_data/base_data_raster')

data = st_as_sf(data, sf_column_name = 'geometry')

data_total = data

data = data[
            (data$admn0Nm == 'Mauritania' & data$admn1Nm %in% c('Hodh Ech Chargi', 'Hodh El Gharbi', 'Tagant', 'Assaba', 'Brakna',    'Gorgol', 'Guidimakha', 'Trarza', 'Nouakchott') ) |
              (data$admn0Nm == 'Senegal' & data$admn1Nm %in% c('Saint Louis', 'Louga', 'Matam', 'Dakar', 'Thies', 'Diourbel') ) |
              (data$admn0Nm == 'Mali' & data$admn1Nm %in% c( 'Tombouctou', 'Gao', 'Kayes', 'Koulikoro', 'Segou', 'Mopti') ) |
              (data$admn0Nm == 'Niger' & data$admn1Nm %in% c('Agadez') == FALSE ) |
              (data$admn0Nm == 'Chad' & data$admn1Nm %in% c('Zone 2','Zone 3','Zone 4','Zone 5') ) |
              (data$admn0Nm == 'Nigeria' & data$admn1Nm %in% c('Sokoto', 'Zamfara','Katsina','Kano','Jigawa','Yobe','Borno') & (data$Wasted > 0 | is.na(data$Wasted))  ) |
              (data$admn0Nm == 'Burkina Faso' & data$admn1Nm %in% c('Sahel', 'Nord', 'Centre-Nord') )
            ,]


data$Bmss_nm = (data$Bmss_nm)

summary(data)

#Remove all columns where wasted is NA
#Remember to change this!
data_wasted = data[!is.na(data$Wasted) & !is.na(data$Bmss_nm) ,]
data_stunted = data[!is.na(data$Stunted) & !is.na(data$Bmss_nm) ,]
data_mrtly = data[!is.na(data$Mortlty) & !is.na(data$Bmss_nm) ,]

#Use this to test the effect of removing the very few points with very high wasting values.
#data = data[data$Wasted <= 30,]

```
## Exploratory analysis

Densities

```{r}

par(mfrow=c(1,3))

hist(na.omit(data$Area), xlab = 'Area', main = 'Density of Area')
hist(na.omit(data$Biomass), xlab = 'Biomass', main = 'Density of Biomass')
hist(na.omit(data$Bmss_nm), xlab = 'Biomass anomaly', main = 'Density of Biomass anomaly')

```

```{r}

par(mfrow=c(1,2))

hist(na.omit(data$Wasted), xlab = 'Wasting prevalence', main = 'Density of wasting prevalence')
hist(na.omit(data$Stunted), xlab = 'Stunting prevalence', main = 'Density of stunting prevalence')
#plot(density(na.omit(data$Mortlty)), xlab = 'Mortality under 5 prevalence', main = 'Density of mortality under 5 prevalence')

```


```{r}

ggplot(data) + aes( y = (Bmss_nm), col = admn0Nm) + geom_boxplot() + ylab('Biomass anomaly')  + theme_classic()

```

```{r}
library(grid)
library(gridExtra)

grid_arrange_shared_legend <- function(...) {
  plots <- list(...)
  g <- ggplotGrob(plots[[1]] + theme(legend.position="bottom"))$grobs
  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
  lheight <- sum(legend$height)
  grid.arrange(
    do.call(arrangeGrob, lapply(plots, function(x)
      x + theme(legend.position="none"))),
    legend,
    ncol = 1,
    heights = unit.c(unit(1, "npc") - lheight, lheight))
}



#plot_mort = ggplot(na.omit(data)) + aes(x = admn0Nm, y = Mortlty, col = admn0Nm) + geom_boxplot() + labs(x = '', fill = 'Country') + theme_classic()
plot_wasted = ggplot(na.omit(data)) + aes(x = admn0Nm, y = Wasted, col = admn0Nm) + geom_boxplot() + labs(x = '', fill = 'Country') + theme_classic()
plot_stunted = ggplot(na.omit(data)) + aes(x = admn0Nm, y = Stunted, col = admn0Nm) + geom_boxplot() + labs(x = '', fill = 'Country') + theme_classic()


grid_arrange_shared_legend(plot_wasted,plot_stunted, ncol=2, top = textGrob("Stunting and wasting prevalences per country",gp=gpar(fontsize=20)))
```

```{r}

x <- aggregate(data_wasted, 
               by = list(data_wasted$admn0Nm,data_wasted$Year),
               FUN = mean,
               warn = FALSE)


x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

wasting_plot = ggplot(data = x) + aes(x = Year, y = Wasted, col = Group.1) + geom_line(size = 1) + geom_point() +  ylab('Wasting prevalence') + labs(fill = 'Country name') + theme_classic()

x <- aggregate(data_stunted, 
               by = list(data_stunted$admn0Nm,data_stunted$Year),
               FUN = mean,
               warn = FALSE)


x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

stunting_plot = ggplot(data = x) + aes(x = Year, y = Stunted, col = Group.1)  + geom_line(size = 1) + geom_point()  + ylab('Stunting prevalence')+ labs(fill = 'Country name') + theme_classic()

grid_arrange_shared_legend(wasting_plot,stunting_plot, ncol=1, top = textGrob("Stunting and wasting prevalences over time",gp=gpar(fontsize=20)))

```


```{r}
library(corrplot)
data_cor = data
data_cor$geometry = NULL
cor1 = cor(na.omit((data_cor[,c(4:6,8:10)])))
corrplot.mixed(cor1, lower.col = 'black', number.cex = .7)

```


```{r}

x <- aggregate(data_wasted, 
               by = list(data_wasted$admn0Nm,data_wasted$admn1Nm),
               FUN = mean,
               warn = FALSE)


x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

x_total <- aggregate(data_wasted, 
               by = list(data_wasted$admn0Nm),
               FUN = mean,
               warn = FALSE)


x_total = st_as_sf(x_total, sf_column_name = 'geometry')
x_total = st_collection_extract(
  x_total,
  type = c("POLYGON"),
  warn = FALSE
)

world_points<- st_centroid(x)
world_points <- cbind(x, st_coordinates(st_centroid(x$geometry)))

world_points_total<- st_centroid(x_total)
world_points_total <- cbind(x_total, st_coordinates(st_centroid(x_total$geometry)))

#This three librarys are necessary to create the map of the world
library(rnaturalearth)
library("rnaturalearthdata")
library(rgeos)
library("ggrepel")
world <- ne_countries(scale = "medium", returnclass = "sf")

theme_set(theme_bw())

```

Map of the Sahel

```{r}

ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data = x, color = 'gray', fill = 'antiquewhite2') +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 3, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y =      c(2, -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('The Sahel') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))

```

Map of West Africa & Sahel

```{r}
# 
# x2 <- aggregate(data_total, 
#                by = list(data_total$admn0Nm,data_total$admn1Nm),
#                FUN = mean,
#                warn = FALSE)
# 
# 
# x2 = st_as_sf(x2, sf_column_name = 'geometry')
# x2 = st_collection_extract(
#   x2,
#   type = c("POLYGON"),
#   warn = FALSE
# )
# 
# x_total <- aggregate(data_total, 
#                by = list(data_total$admn0Nm),
#                FUN = mean,
#                warn = FALSE)
# 
# 
# x_total = st_as_sf(x_total, sf_column_name = 'geometry')
# x_total = st_collection_extract(
#   x_total,
#   type = c("POLYGON"),
#   warn = FALSE
# )
# 
# world_points<- st_centroid(x2)
# world_points <- cbind(x2, st_coordinates(st_centroid(x2$geometry)))
# 
# world_points_total<- st_centroid(x_total)
# world_points_total <- cbind(x_total, st_coordinates(st_centroid(x_total$geometry)))
# 
# ggplot(data = world) +
#     geom_sf(fill = "antiquewhite1") +
#     geom_sf(data = x2, color = 'darkgray', fill = 'antiquewhite3') +
#     geom_sf(data = x, color = 'gray', fill = 'antiquewhite2') +
#     geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
#     geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 3, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y =      c(2, -0.25, 1.5, 1.5, -0.5)) +
#     scale_fill_viridis_c(alpha = .5) +
#     geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
#     coord_sf(xlim =c(30, -20), ylim = c(0, 28), expand = FALSE) + 
#     xlab('Longitude') + ylab('Latitude') + 
#     ggtitle('West Africa & The Sahel', subtitle = 'The sahel ') + 
#     theme( panel.background = element_rect(fill = 'aliceblue'))

```

Wasting

```{r errors=FALSE}




biomass_raster = raster(x, res = 1/100)
rasterize_biomass = fasterize(x, biomass_raster, field = 'Wasted')


ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data = x, aes(fill = x$Wasted), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean acute malnutrition prevalence in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Wasting prevalence') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))
```

Stunting

```{r}


x <- aggregate(data_stunted, 
               by = list(data_stunted$admn0Nm,data_stunted$admn1Nm),
               FUN = mean,
               warn = FALSE)


x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

x_total <- aggregate(data_stunted, 
               by = list(data_stunted$admn0Nm),
               FUN = mean,
               warn = FALSE)


x_total = st_as_sf(x_total, sf_column_name = 'geometry')
x_total = st_collection_extract(
  x_total,
  type = c("POLYGON"),
  warn = FALSE
)

biomass_raster = raster(x, res = 1/100)
rasterize_biomass = fasterize(x, biomass_raster, field = 'Stunted')


ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data = x, aes(fill = x$Stunted), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean chronic malnutrition prevalence in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Stunting prevalence') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))
```

Mortality

```{r}


x <- aggregate(data_mrtly, 
               by = list(data_mrtly$admn0Nm,data_mrtly$admn1Nm),
               FUN = mean,
               warn = FALSE)


x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

x_total <- aggregate(data_mrtly, 
               by = list(data_mrtly$admn0Nm),
               FUN = mean,
               warn = FALSE)


x_total = st_as_sf(x_total, sf_column_name = 'geometry')
x_total = st_collection_extract(
  x_total,
  type = c("POLYGON"),
  warn = FALSE
)

biomass_raster = raster(x, res = 1/100)
rasterize_biomass = fasterize(x, biomass_raster, field = 'Mortlty')


ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data = x, aes(fill = x$Mortlty), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean mortality under five prevalence in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Mortality prevalence') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))
```


Biomass

```{r}

x <- aggregate(data, 
               by = list(data$admn0Nm,data$admn1Nm),
               FUN = mean,
               warn = FALSE)


x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

x_total <- aggregate(data, 
               by = list(data$admn0Nm),
               FUN = mean,
               warn = FALSE)


x_total = st_as_sf(x_total, sf_column_name = 'geometry')
x_total = st_collection_extract(
  x_total,
  type = c("POLYGON"),
  warn = FALSE
)


ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data = x, aes(fill = x$Bmss_nm), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 3, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean biomass anomaly in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Biomass anomaly') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))


```

## Modeling Wasting


#Linear regression

We are going to star by using a simple model. There we are going to plot the residuals in the raster to see if they seem to be distributed as random.

We are using the best model as determined by the mixed model basic.

```{r}

model.lineal = lm(Wasted ~  Year + Area + Bmss_nm, data = data_wasted)
summary(model.lineal)

```

```{r}

plot(model.lineal)

```

```{r}

model.lineal.2 = lm(log(Wasted) ~  Year + Area + Bmss_nm, data = data_wasted)
summary(model.lineal.2)

```

```{r}

plot(model.lineal.2)

```


```{r}

model.lineal.3 = lm(sqrt(Wasted) ~  Year + Area + Bmss_nm, data = data_wasted)
summary(model.lineal.3)

```

```{r}

plot(model.lineal.3)

```

No parecen tener mucho efecto las transformaciones.

```{r}

data_wasted$residuals = residuals(model.lineal)

x <- aggregate(data_wasted, 
               by = list(data_wasted$admn0Nm,data_wasted$admn1Nm),
               FUN = mean)

x
x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data =x, aes(fill = x$residuals), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean residuals of estimated acute malnutrition prevalence in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Estimated residuals') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))



```

Residuals per year

```{r}

for(i in 1998:2018){
  
  ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data =data_wasted[data_wasted$Year == i,], aes(fill = data_wasted[data_wasted$Year == i,]$residuals), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean acute malnutrition prevalence in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Wasting rate') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))
}




```

# Mixed models

How to choose random effects and fixed effects?

We start by testing random effects, then, the fixed effects.


```{r}

#Random intercept with fixed mean
model.mixed.1 =lmer(Wasted ~ Year + Area + Bmss_nm + (1|admn0Nm) + (1|admn1Nm), data = data_wasted, REML = FALSE)
#Intercept varying among group 1 and group 2 within group 1
model.mixed.2 = lmer(Wasted ~ Year + Area + Bmss_nm + (1|admn0Nm/admn0Nm), data = data_wasted, REML = FALSE)
summary(model.mixed.1)
summary(model.mixed.1)


```

```{r}

test12=-2*logLik(model.mixed.1, REML=T) +2*logLik(model.mixed.2, REML=T)
mean(pchisq(test12,df=c(0,1),lower.tail=F))
#anova(model.mixed.1,model.mixed.2)

```

As we can see, independent intercepts for each country and region seem to perform better than nested as shown by ANOVA.

```{r}
#Uncorrelated slope and intercept for the year and intercept varying among g1 and g2 withing g1.
model.mixed.3 = lmer(Wasted ~ Year + Area + Bmss_nm + (Year  | admn0Nm) + (Year  | admn1Nm), data = data_wasted, REML = FALSE, na.action = na.omit)
summary(model.mixed.3)
anova(model.mixed.1,model.mixed.3)
test13=-2*logLik(model.mixed.1, REML=T) +2*logLik(model.mixed.3, REML=T)
mean(pchisq(test13,df=c(0,1),lower.tail=F))

```

Doesnt seem to improve the fit

```{r}

#Uncorrelated slope and intercept for the year and intercept varying among g1 and g2 withing g1.
model.mixed.4 = update(model.mixed.1, .~.+Year:Bmss_nm+Area:Bmss_nm)
model.mixed.5 = update(model.mixed.2, .~.+Year:Bmss_nm+Area:Bmss_nm)

anova(model.mixed.4,model.mixed.5)
test45=-2*logLik(model.mixed.4, REML=T) +2*logLik(model.mixed.5, REML=T)
mean(pchisq(test45,df=c(0,1),lower.tail=F))

```

```{r}

anova(model.mixed.4,model.mixed.1)

```

As we can see, the best model seem to be the model number 1, using intercepts varying among admn0Nm and admn1Nm (independent intercepts).

We inspect model 1 now:

```{r}

model.mixed.1 = update(model.mixed.1, REML = TRUE)

data_wasted$residuals = residuals(model.mixed.1)

x <- aggregate(data_wasted, 
               by = list(data_wasted$admn0Nm,data_wasted$admn1Nm),
               FUN = mean)

x
x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data =x, aes(fill = x$residuals), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean residuals of estimated acute malnutrition prevalence in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Estimated residuals') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))



```

Residuals per year

```{r}

for(i in 1998:2018){
  
  ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data =data_wasted[data_wasted$Year == i,], aes(fill = data_wasted[data_wasted$Year == i,]$residuals), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean residuals of estimated acute malnutrition prevalence in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Estimated residuals') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))
}


```

Diagnostic plots to see if the assumptions of the regression are being held.


```{r}

plot(model.mixed.1, form = resid(.) ~ Year | admn1Nm)

```


```{r}

plot(model.mixed.1)
par(mfrow=c(1,2))
plot(model.mixed.1)
qqnorm(resid(model.mixed.1))
qqline(resid(model.mixed.1))


```
Variance does not seem to be constant:


```{r}

plot(model.mixed.1 , Wasted ~ fitted(.) , abline = c(0,1))
plot(model.mixed.1 , Wasted ~ fitted(.) | admn0Nm, abline = c(0,1))
plot(model.mixed.1 , Wasted ~ fitted(.) | admn1Nm, abline = c(0,1))

```

There seems to be spatial autocorrelation, we test that with Moran's I

```{r}

nb <- poly2nb(data_wasted)
lw <- nb2listw(nb)

moran.test(data_wasted$Wasted, lw, randomisation = TRUE)
moran.mc(data_wasted$Wasted, lw, nsim = 10000)
```

The p-value lead us to reject the null hypothesis, as a result, there is strong evidence that the data is not independently normally distributed and I > 0 indicating spatial correlation. As a result, we fit the spatial model.

```{r}

library(spaMM)
library(spatialreg)
library(spdep)
library(proj4)

```

First the model without spatial components:

```{r}

model.spam = fitme(Wasted ~ Year + Area + Bmss_nm + (1 | admn0Nm) + (1 | admn1Nm) ,data=data_wasted, method = 'ML')

summary(model.spam)

```


```{r}

#data = data[data$Year == 2018,]

options(warn=-1)




#unique_data = data[unique(data$admn1Nm),]

centroids <- st_transform(data_wasted$geometry, 29101) %>% 
  st_centroid() %>% 
  # this is the crs from d, which has no EPSG code:
  st_transform(., '+proj=longlat +ellps=GRS80 +no_defs') %>%
  # since you want the centroids in a second geometry col:
  st_geometry()

# check with
plot(st_geometry(centroids))
plot(centroids[, 'centroids'], add = T, col = 'red', pch = 19)

coordinates = st_coordinates(centroids)

data_wasted$lat = coordinates[,1]
data_wasted$lon = coordinates[,2]

proj4string <- "+proj=utm +zone=19 +south +ellps=WGS84 +datum=WGS84 +units=m +no_defs "
pj <- project(centroids, proj4string, inverse=TRUE)

library(surveillance)
adj = (poly2adjmat(data_wasted))

MLdistMat <- as.matrix(proxy::dist(centroids))

```

```{r}
data_wasted.2 = data_wasted
data_wasted.2$geometry = NULL
model.spam.2 = fitme(Wasted ~ Year + Area + Bmss_nm + (1 | admn0Nm) + (1 | admn1Nm) +  Matern( 1 | lat + lon), distMatrix = MLdistMat, data=data_wasted.2, method = 'ML')
summary(model.spam.2)
```


```{r}
#testing bubble
# data_wasted.2$Bmss_nm = data_wasted.2$Bmss_nm/10000
# model.spam.3 = fitme(Wasted ~ Year + Area + Bmss_nm + (1 | admn0Nm) + (1 | admn1Nm) +  Matern( 1 | lat + lon) + AR1(1| Year ), distMatrix = MLdistMat, data=data_wasted.2, method = 'ML')
# 
# summary(model.spam.3)

test12=-2*logLik(model.spam, REML=T) +2*logLik(model.spam.2, REML=T)
mean(pchisq(test12,df=c(0,1),lower.tail=F))

```


```{r}

model.spam.2 = update(model.spam.2, method = 'REML')

data_wasted$residuals = residuals(model.spam.2)

x <- aggregate(data_wasted, 
               by = list(data_wasted$admn0Nm,data_wasted$admn1Nm),
               FUN = mean)

x
x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data =x, aes(fill = x$residuals), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean residuals of estimated acute malnutrition prevalence in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Estimated residuals') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))



```

```{r}

plot(model.spam.2, form = resid(.) ~ Year | admn1Nm)

```


```{r}

plot(model.spam.2)
par(mfrow=c(1,2))
plot(model.spam.2)
qqnorm(resid(model.spam.2))
qqline(resid(model.spam.2))


```
Variance does not seem to be constant:


```{r}

plot(model.spam.2$eta, data_wasted$Wasted)
abline(a = 0,b = 1)

```

```{r}

if (suppressPackageStartupMessages(require(maps,quietly=TRUE))) {
## 'maps' required for add.map=TRUE
filled.mapMM(model.spam.2,add.map=TRUE,plot.axes=quote({axis(1);axis(2)}),
decorations=quote(points(pred[,coordinates],pch=15,cex=0.3)),
plot.title=title(main="Inferred prevalence of acute malnutrition, Sahel",
xlab="longitude",ylab="latitude"), yrange = c(5,27))
}


```









## Modeling Stunting

#Linear regression

We are going to star by using a simple model. There we are going to plot the residuals in the raster to see if they seem to be distributed as random.


```{r}

model.lineal = lm(Stunted ~  Year + Area + Bmss_nm, data = data_stunted)
summary(model.lineal)

```

```{r}

plot(model.lineal)

```

```{r}

model.lineal.2 = lm(log(Stunted) ~  Year + Area + Bmss_nm, data = data_stunted)
summary(model.lineal.2)

```

```{r}

plot(model.lineal.2)

```


```{r}

model.lineal.3 = lm(sqrt(Stunted) ~  Year + Area + Bmss_nm, data = data_stunted)
summary(model.lineal.3)

```

```{r}

plot(model.lineal.3)

```

No parecen tener mucho efecto las transformaciones.

```{r}

data_stunted$residuals = residuals(model.lineal)

x <- aggregate(data_stunted, 
               by = list(data_stunted$admn0Nm,data_stunted$admn1Nm),
               FUN = mean)

x
x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data =x, aes(fill = x$residuals), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean residuals of estimated chronic malnutrition prevalence in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Estimated residuals') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))



```

Residuals per year

```{r}

for(i in 1998:2018){
  
  ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data =data_wasted[data_wasted$Year == i,], aes(fill = data_wasted[data_wasted$Year == i,]$residuals), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Residuals of estimated chronic malnutrition prevalence in the Sahel', subtitle = 'Year ' + i + ', by region') + 
    labs(fill = 'Wasting rate') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))
}




```

# Mixed models

How to choose random effects and fixed effects?

We start by testing random effects, then, the fixed effects.


```{r}

#Random intercept with fixed mean
model.mixed.1 =lmer(Stunted ~ Year + Area + Bmss_nm + (1|admn0Nm) + (1|admn1Nm), data = data_stunted, REML = FALSE)
#Intercept varying among group 1 and group 2 within group 1
model.mixed.2 = lmer(Stunted ~ Year + Area + Bmss_nm + (1|admn0Nm/admn0Nm), data = data_stunted, REML = FALSE)
summary(model.mixed.1)
summary(model.mixed.1)


```

```{r}

anova(model.mixed.1,model.mixed.2)

```

As we can see, independent intercepts for each country and region seem to perform better than nested as shown by ANOVA.

```{r}
#Uncorrelated slope and intercept for the year and intercept varying among g1 and g2 withing g1.
model.mixed.3 = lmer(Stunted ~ Year + Area + Bmss_nm + (Year  | admn0Nm) + (Year  | admn1Nm), data = data_stunted, REML = FALSE, na.action = na.omit)
summary(model.mixed.3)
anova(model.mixed.1,model.mixed.3)

```

Doesnt seem to improve the fit

```{r}

#Uncorrelated slope and intercept for the year and intercept varying among g1 and g2 withing g1.
model.mixed.4 = update(model.mixed.1, .~.+Year:Bmss_nm+Area:Bmss_nm)
model.mixed.5 = update(model.mixed.2, .~.+Year:Bmss_nm+Area:Bmss_nm)

anova(model.mixed.4,model.mixed.5)
```

```{r}

anova(model.mixed.4,model.mixed.1)

```

As we can see, the best model seem to be the model number 1, using intercepts varying among admn0Nm and admn1Nm (independent intercepts).

We inspect model 1 now:

```{r}

model.mixed.1 = update(model.mixed.1, REML = TRUE)

data_stunted$residuals = residuals(model.mixed.1)

x <- aggregate(data_stunted, 
               by = list(data_stunted$admn0Nm,data_stunted$admn1Nm),
               FUN = mean)

x
x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data =x, aes(fill = x$residuals), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean residuals of estimated chronic malnutrition prevalence in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Estimated residuals') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))



```

Residuals per year

```{r}

for(i in 1998:2018){
  
  ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data =data_wasted[data_wasted$Year == i,], aes(fill = data_wasted[data_wasted$Year == i,]$residuals), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Residuals of estimated chronic malnutrition prevalence in the Sahel', subtitle = 'Year ' + i + ', by region') + 
    labs(fill = 'Estimated residuals') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))
}


```

Diagnostic plots to see if the assumptions of the regression are being held.


```{r}

plot(model.mixed.1, form = resid(.) ~ Year | admn1Nm)

```


```{r}

plot(model.mixed.1)
par(mfrow=c(1,2))
plot(model.mixed.1)
qqnorm(resid(model.mixed.1))
qqline(resid(model.mixed.1))


```
Variance does not seem to be constant:


```{r}

plot(model.mixed.1 , Stunted ~ fitted(.) , abline = c(0,1))
plot(model.mixed.1 , Stunted ~ fitted(.) | admn0Nm, abline = c(0,1))
plot(model.mixed.1 , Stunted ~ fitted(.) | admn1Nm, abline = c(0,1))

```

There seems to be spatial autocorrelation, we test that with Moran's I

```{r}

nb <- poly2nb(data_stunted)
lw <- nb2listw(nb)

moran.test(data_stunted$Stunted, lw, randomisation = TRUE)
moran.mc(data_stunted$Stunted, lw, nsim = 10000)
```

The p-value lead us to reject the null hypothesis, as a result, there is strong evidence that the data is not independently normally distributed and I > 0 indicating spatial correlation. As a result, we fit the spatial model.


First the model without spatial components:

```{r}

model.spam = fitme(Stunted ~ Year + Area + Bmss_nm + (1 | admn0Nm) + (1 | admn1Nm) ,data=data_stunted, method = 'ML')

summary(model.spam)

```


```{r}

#data = data[data$Year == 2018,]

options(warn=-1)


#unique_data = data[unique(data$admn1Nm),]

centroids <- st_transform(data_stunted$geometry, 29101) %>% 
  st_centroid() %>% 
  # this is the crs from d, which has no EPSG code:
  st_transform(., '+proj=longlat +ellps=GRS80 +no_defs') %>%
  # since you want the centroids in a second geometry col:
  st_geometry()

# check with
plot(st_geometry(centroids))
plot(centroids[, 'centroids'], add = T, col = 'red', pch = 19)

coordinates = st_coordinates(centroids)

data_stunted$lat = coordinates[,1]
data_stunted$lon = coordinates[,2]

proj4string <- "+proj=utm +zone=19 +south +ellps=WGS84 +datum=WGS84 +units=m +no_defs "
pj <- project(centroids, proj4string, inverse=TRUE)

library(surveillance)
adj = (poly2adjmat(data_stunted))

MLdistMat <- as.matrix(proxy::dist(centroids))

```

```{r}
data_stunted.2 = data_stunted
data_stunted.2$geometry = NULL
model.spam.2 = fitme(Stunted ~ Year + Area + Bmss_nm + (1 | admn0Nm) + (1 | admn1Nm) +  Matern( 1 | lat + lon), distMatrix = MLdistMat, data=data_stunted.2, method = 'ML')
summary(model.spam.2)
```


```{r}

model.spam.2 = update(model.spam.2, method = 'REML')

data_stunted$residuals = residuals(model.spam.2)

x <- aggregate(data_stunted, 
               by = list(data_stunted$admn0Nm,data_stunted$admn1Nm),
               FUN = mean)

x
x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data =x, aes(fill = x$residuals), color = NA) +
    geom_sf(data = world, fill = NA, size = 0.6, color = 'black') +
    geom_label_repel(data = world_points_total, aes(X, Y, label = Group.1), size = 4, fontface = "bold", alpha = 0.6, nudge_x = c(3, -2, 3, 2, -1), nudge_y = c(2, 
    -0.25, 1.5, 1.5, -0.5)) +
    scale_fill_viridis_c(alpha = .5) +
    geom_text_repel(data= world_points,aes(x=X, y=Y, label=Group.2), color = 'black', fontface = 'bold', check_overlap = TRUE,size = 3) +
    coord_sf(xlim =c(25, -20), ylim = c(5, 28), expand = FALSE) + 
    xlab('Longitude') + ylab('Latitude') + 
    ggtitle('Mean residuals of estimated chronic malnutrition prevalence in the Sahel', subtitle = 'From 1998 to 2018, by region') + 
    labs(fill = 'Estimated residuals') + 
    theme( panel.background = element_rect(fill = 'aliceblue'))



```

```{r}

plot(model.spam.2, form = resid(.) ~ Year | admn1Nm)

```


```{r}

plot(model.spam.2)
par(mfrow=c(1,2))
plot(model.spam.2)
qqnorm(resid(model.spam.2))
qqline(resid(model.spam.2))


```
Variance does not seem to be constant:


```{r}

plot(model.spam.2$eta, data_stunted$Stunted)
abline(a = 0,b = 1)

```

```{r}

if (suppressPackageStartupMessages(require(maps,quietly=TRUE))) {
## 'maps' required for add.map=TRUE
filled.mapMM(model.spam.2,add.map=TRUE,plot.axes=quote({axis(1);axis(2)}),
decorations=quote(points(pred[,coordinates],pch=15,cex=0.3)),
plot.title=title(main="Inferred prevalence of chronic malnutrition, Sahel",
xlab="longitude",ylab="latitude"), yrange = c(5,27))
}


```





