---
title: "Mixed model spatial"
author: "Ignacio García Sánchez-Migallón"
date: "17/07/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('SASmixed')
library(nlme)
library(lme4)
library(ggplot2)
library(grid)
library(gridExtra)
library(gstat)
library(tidyverse)
library(fasterize)
library(tiff)
library(raster)
library(sf)
rm(list=ls())

```

## Introduction

Perform an exploratory analysis of the data.
All the series of plots generated with grid. have to be done in the console as to obtain bigger plots.

```{r}

setwd("C:/Users/IgnacioGarcia/Desktop/UNIVERSIDAD/Statistics for data science/Master thesis/data")
data = st_read('Mixed_model_data/base_data_raster')
#data$X = NULL

data = st_as_sf(data, sf_column_name = 'geometry')


#data$Biomass_anomaly = bestNormalize(data$Biomass_anomaly)$x.t

#data$Biomass_anomaly = scale(data$Biomass_anomaly)
data$Bmss_nm = (data$Bmss_nm)
data$Area = (data$Area)

summary(data)

#Remove all columns where wasted is NA
#data = data[!is.na(data$Stunted) & !is.na(data$Bmss_nm) ,]

```


## Modeling Wastedness

```{r}

# data = data %>%
#   group_by(admn0Nm, admn1Nm) %>%
#   summarise( Wasted = mean(Wasted), geometry = merge(geometry, geometry))
x <- aggregate(data, 
               by = list(data$admn0Nm,data$admn1Nm),
               FUN = mean)

x
x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

biomass_raster = raster(x, res = 1/100)
rasterize_biomass = fasterize(x, biomass_raster, field = 'Wasted')
plot(rasterize_biomass, main = 'Mean value per admin1')


```

Just to get a nice visualization of the area:

```{r}

world_points<- st_centroid(data)
world_points <- cbind(data, st_coordinates(st_centroid(data$geometry)))

ggplot(data = data) +geom_sf(data = x, size = 0.25, color = "black",aes(fill= x$Stunted)) + geom_text(data= world_points,aes(x=X, y=Y, label=admn1Nm), color = 'burlywood3', fontface = 'bold', check_overlap = FALSE,size = 2.5) + annotate(geom = 'text', x = -90, y = 26, label = 'Sahel Map', fontface = 'italic', color = 'grey22', size = 6, location = 'bl', width_hint = 0.5)  + coord_sf(xlim = c(25, -20), ylim = c(0, 28), expand = FALSE) + xlab('Longitude') + ylab('Latitude') + ggtitle('Mean chronic malnutrition rates in West Africa') + labs(fill = 'Stunting rate') + theme(panel.grid.major = element_line(color = gray(.5), linetype = 'dashed', size = 0.5), panel.background = element_rect(fill = 'aliceblue'))


```


We are going to star by using a simple model, result of the previous basic model analysis. There we are going to plot the residuals in the raster to see if they seem to be distributed as random.

We are using the best model as determined by the mixed model basic.

```{r}

model.1 = lme(Wasted ~ Year + Area + Bmss_nm + Area:Bmss_nm, random = list(admn1Nm = pdDiag(~admn0Nm)), data = data, na.action = na.omit)
summary(model.1, corr = FALSE)


```

```{r}

residuals  = residuals(model.1)

```

```{r}

data$residuals = residuals

x <- aggregate(data, 
               by = list(data$admn0Nm,data$admn1Nm),
               FUN = mean)

x
x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)

biomass_raster = raster(x, res = 1/100)
rasterize_biomass = fasterize(x, biomass_raster, field = 'residuals')
plot(rasterize_biomass, main = 'Mean residuals per admin1')



```
Are the residuals correlated spatially?

```{r}

library(spatialreg)


nb <- poly2nb(data)
lw <- nb2listw(nb)


lag = lagsarlm(residuals ~ 1, data=data, lw, tol.solve=1.0e-30)

summary(lag)
  
  
```


```{r}

x = data[data$Year == 2018,]

biomass_raster = raster(x, res = 1/100)
rasterize_biomass = fasterize(x, biomass_raster, field = 'residuals')
plot(rasterize_biomass, main = 'Mean residuals value per admin1 in 2018')



```

There seems to be certain relationship among the residuals as seen in the raster, the patron does not seem to be random, which may imply certain spatial correlation.

Now we are going to try an Spatial Lag Model.


```{r}

spatial.1 = lagsarlm(Wasted ~ Year + Area + Bmss_nm + Area:Bmss_nm, data=data, lw, tol.solve=1.0e-30)
summary(spatial.1)

```

```{r}

data$residuals <- residuals(spatial.1)


x <- aggregate(data, 
               by = list(data$admn0Nm,data$admn1Nm),
               FUN = mean)

x
x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)
```

```{r}
biomass_raster = raster(x, res = 1/100)
rasterize_biomass = fasterize(x, biomass_raster, field = 'residuals')
plot(rasterize_biomass, main = 'Mean residuals value per admin1')

```


Spatial error model:

```{r}
library(spatialreg)


#nb <- poly2nb(data)
#lw <- nb2listw(nb)

spatial.2 = errorsarlm(Wasted ~ Year + Area + Bmss_nm + Area:Bmss_nm, data=data, lw, tol.solve=1.0e-30)
summary(spatial.2)

```

```{r}

data$residuals <- residuals(spatial.2)

x <- aggregate(data, 
               by = list(data$admn0Nm,data$admn1Nm),
               FUN = mean)

x
x = st_as_sf(x, sf_column_name = 'geometry')
x = st_collection_extract(
  x,
  type = c("POLYGON"),
  warn = FALSE
)
```

```{r}

biomass_raster = raster(x, res = 1/100)
rasterize_biomass = fasterize(x, biomass_raster, field = 'residuals')
plot(rasterize_biomass, main = 'Mean residuals value per admin1')


```



```{r}

BIC(model.1)
BIC(spatial.1)
BIC(spatial.2)

```


Need to try this library:mgcv for GAM's where apparently I can add random effects + the spatial autocorrelation.


