---
title: "Code testing"
author: "Ignacio García Sánchez-Migallón"
date: "08/06/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('SASmixed')
library(nlme)
library(lme4)
library(ggplot2)
library(grid)
library(gridExtra)
rm(list=ls())

```

## Introduction

Perform an exploratory analysis of the data.
All the series of plots generated with grid. have to be done in the console as to obtain bigger plots.

```{r}

setwd("C:/Users/IgnacioGarcia/Desktop/UNIVERSIDAD/Statistics for data science/Master thesis/data")
data = read.csv('Mixed_model_data/base_data.csv')
data$X = NULL


#data$Biomass_anomaly = bestNormalize(data$Biomass_anomaly)$x.t

#data$Biomass_anomaly = scale(data$Biomass_anomaly)
data$Biomass_anomaly = (data$Biomass_anomaly)/1000
data$Area = (data$Area)

summary(data)

```

Theres few missing values for all explanatory variables, response ones have way too many. But thats what we have to deal with.

```{r}

ggplot(data) + aes(x = Year, y = Biomass, col = admin0Name) + geom_point() + theme_classic()

```

```{r}

ggplot(data) + aes( y = (Biomass), col = admin0Name) + geom_boxplot() + theme_classic()

```

```{r}

ggplot(data) + aes(x = Area, y = Biomass,col = admin0Name,  fill = admin0Name) + geom_col(size = 1) + theme_classic()

```

```{r}


grid_arrange_shared_legend <- function(...) {
  plots <- list(...)
  g <- ggplotGrob(plots[[1]] + theme(legend.position="bottom"))$grobs
  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
  lheight <- sum(legend$height)
  grid.arrange(
    do.call(arrangeGrob, lapply(plots, function(x)
      x + theme(legend.position="none"))),
    legend,
    ncol = 1,
    heights = unit.c(unit(1, "npc") - lheight, lheight))
}



plot_mort = ggplot(na.omit(data)) + aes(x = admin0Name, y = Mortality, col = admin0Name) + geom_boxplot() + labs(x = '', fill = 'Country') + theme_classic()
plot_wasted = ggplot(na.omit(data)) + aes(x = admin0Name, y = Wasted, col = admin0Name) + geom_boxplot() + labs(x = '', fill = 'Country') + theme_classic()
plot_stunted = ggplot(na.omit(data)) + aes(x = admin0Name, y = Stunted, col = admin0Name) + geom_boxplot() + labs(x = '', fill = 'Country') + theme_classic()


grid_arrange_shared_legend(plot_mort,plot_wasted,plot_stunted, ncol=2, top = textGrob("Mortality, Stunted and Wasted rates per country",gp=gpar(fontsize=20)))
```


```{r}

plot_mort = ggplot(na.omit(data)) + aes(x = Year, y = Mortality, col = admin0Name)+ geom_point(size = 0.5) + geom_quantile(size = 1.5, quantiles = 0.5) + theme_classic()
plot_wasted = ggplot(na.omit(data)) + aes(x = Year, y = Wasted, col = admin0Name)+ geom_point(size = 0.5) + geom_quantile(size = 1.5, quantiles = 0.5) + theme_classic()
plot_stunted = ggplot(na.omit(data)) + aes(x = Year, y = Stunted, col = admin0Name)+ geom_point(size = 0.5) + geom_quantile(size = 1.5, quantiles = 0.5) + theme_classic()

grid_arrange_shared_legend(plot_mort,plot_wasted,plot_stunted, nrow=2, top = textGrob("Mortality, Stunted and Wasted rates per country",gp=gpar(fontsize=20)))

```

```{r}
#Corrplot
library(corrplot)
cor1 = cor(na.omit(data[,3:10]))
corrplot.mixed(cor1, lower.col = 'black', number.cex = .7)

```

In the corrplot it is easy to notice how the admins have dissapeared, as they are not numerical they cannot be shown in this plot.


## Modeling Wastedness

Justification about fixed effects and random effects.

```{r}

model.1 = lme(Wasted ~ Year  + Area + Biomass_anomaly , random = ~ 1 | admin1Name, data = data, na.action = na.omit)
summary(model.1, corr = FALSE)



```

```{r}

fixef(model.1)

```

```{r}

model.0 = lm(Wasted ~ Year + Mean + Biomass_anomaly + Area, data = data, na.action = na.omit)
test=-2*logLik(model.0) +2*logLik(model.1)
mean(pchisq(test,df=c(0,1),lower.tail=F))
```

This shows that the random effect is necessary.

```{r}

model.2 = lme(Wasted ~ Year + Area + Biomass_anomaly, random = list(admin1Name = pdDiag(~admin0Name)), data = data, na.action = na.omit)
summary(model.2, corr = FALSE)




```

```{r}

fixef(model.2)

```


```{r}

model.3 = lme(Wasted ~ Year + Area + Biomass_anomaly, random = ~ 1 | admin0Name/admin1Name, data = data, na.action = na.omit)
summary(model.3, corr = FALSE)


```


```{r}

fixef(model.3)

```


```{r}

model.4 = update(model.2, .~.+Year:Biomass_anomaly+Area:Biomass_anomaly)
model.5 = update(model.3, .~.+Year:Biomass_anomaly+Area:Biomass_anomaly)


```

```{r}

anova(model.4,model.5)

```

```{r}
#Now model 4 versus model 1.
model.1b = update(model.1, .~.+Year:Biomass_anomaly+Area:Biomass_anomaly)
test=-2*logLik(model.1b) +2*logLik(model.4)
mean(pchisq(test,df=c(0,1),lower.tail=F))

```

```{r}
model.4 = update(model.4, method = 'ML')
model.2 = update(model.2, method = 'ML')
anova(model.2,model.4)

```

```{r}
#Model final?
model.4 = update(model.4, method = 'REML')
summary(model.4)
```

```{r}

model.4 = update(model.4, .~.-Year:Biomass_anomaly, method = 'REML')
summary(model.4)

```

```{r}

fixef(model.4)

```

```{r}

plot(model.4 , Wasted ~ fitted(.) | admin0Name, abline = c(0,1))
plot(model.4 , Wasted ~ fitted(.) | admin1Name, abline = c(0,1))

```

```{r}
#This panel is for testing exclusively.

# Third possibility
plot(model.4 , Wasted ~ Year | admin1Name, abline = c(0,1), ylim = c(0,50)) 
plot(model.4 , fitted(.) ~ Year | admin1Name, abline = c(0,1), ylim = c(0,50))

```


```{r}

plot(model.4)

```
Not constant?

```{r}

model.4b=update(model.4,weights=varPower(form=~fitted(.)))

plot(model.4b)

```


```{r}
par(mfrow=c(1,2))
plot(model.4b)
qqnorm(resid(model.4b))
qqline(resid(model.4b))
shapiro.test(resid(model.4b))


```

We can see that the residuals are not normal. This might be solved by transforming the response variable Wasted. Let's try with a log transformation.

```{r}
library(boxcoxmix)
library(bestNormalize)


data_alt = data
#data_alt$Wasted = bestNormalize(data$Wasted)$x.t

model.wasting = lme(Wasted ~ Year + Area +  Biomass_anomaly + Area:Biomass_anomaly, random = list(admin1Name = pdDiag(~admin0Name)), data = data_alt,na.action = na.omit )


```

```{r}

summary(model.wasting)
plot(model.wasting)
qqnorm(resid(model.wasting))
qqline(resid(model.wasting))
shapiro.test(resid(model.wasting))


```

```{r}

plot(model.wasting , Wasted ~ fitted(.) | admin0Name, abline = c(0,1))
plot(model.wasting , Wasted ~ fitted(.) | admin1Name, abline = c(0,1))

```


```{r}

plot(ranef(model.4b))

```


## Modeling Stuntedness

Justification about fixed effects and random effects.

```{r}

model.1 = lme(Stunted ~ Year + Biomass_anomaly + Area, random = ~ 1 | admin1Name, data = data, na.action = na.omit)
summary(model.1, corr = FALSE)



```

```{r}

fixef(model.1)

```

```{r}

model.0 = lm(Stunted ~ Year + Mean + Biomass + Area, data = data, na.action = na.omit)
test=-2*logLik(model.0) +2*logLik(model.1)
mean(pchisq(test,df=c(0,1),lower.tail=F))
```

This shows that the random effect is necessary.

```{r}

model.2 = lme(Stunted ~ Year  + Biomass_anomaly + Area, random = list(admin1Name = pdDiag(~admin0Name)), data = data, na.action = na.omit)
summary(model.2, corr = FALSE)




```

```{r}

fixef(model.2)

```


```{r}

model.3 = lme(Stunted ~ Year + Biomass_anomaly + Area, random = ~ 1 | admin0Name/admin1Name, data = data, na.action = na.omit)
summary(model.3, corr = FALSE)


```


```{r}

fixef(model.3)

```


```{r}

model.4 = update(model.2, .~.+Year:Biomass_anomaly+Area:Biomass_anomaly)
model.5 = update(model.3, .~.+Year:Biomass_anomaly+Area:Biomass_anomaly)


```

```{r}

anova(model.4,model.5)

```

```{r}
#Now model 5 versus model 1.
model.1b = update(model.1, .~.+Year:Biomass_anomaly+Area:Biomass_anomaly)
test=-2*logLik(model.1b) +2*logLik(model.5)
mean(pchisq(test,df=c(0,1),lower.tail=F))

```

```{r}
model.5 = update(model.5, method = 'ML')
model.3 = update(model.3, method = 'ML')
anova(model.3,model.5)

```

```{r}
#Model final?
model.3 = update(model.3, method = 'REML')
summary(model.3)
```

```{r}
model.3 = update(model.3,.~.-Biomass_anomaly, method = 'REML')
summary(model.3)
fixef(model.3)

```

```{r}

plot(model.3 , Stunted ~ fitted(.) | admin0Name, abline = c(0,1))
plot(model.3 , Stunted ~ fitted(.) | admin1Name, abline = c(0,1))

```

```{r}

plot(model.3)

```
Not constant?

```{r}

model.3b=update(model.3,weights=varPower(form=~fitted(.)))

plot(model.3)

```


```{r}

qqnorm(resid(model.3b))
qqline(resid(model.3b))
shapiro.test(resid(model.3b))


```

Again, we are not respecting the heterocedasticity of the regression, so we need to transform the response variable Stunting.

```{r}

#data$Stunted = sqrt_x(data$Stunted)$x.t

model.final = lme(Stunted ~ Year  + Area, random = ~ 1 | admin0Name/admin1Name, data = data, na.action = na.omit)


summary(model.final)
plot(model.final)
qqnorm(resid(model.final))
qqline(resid(model.final))
shapiro.test(resid(model.final))


```

```{r}

plot(model.final , Stunted ~ fitted(.) | admin0Name, abline = c(0,1))
plot(model.final , Stunted ~ fitted(.) | admin1Name, abline = c(0,1))

```





```{r}

plot(ranef(model.3))

```

## Modeling mortality

Justification about fixed effects and random effects.

```{r}

model.1 = lme(Mortality ~ Year  + Biomass_anomaly + Area, random = ~ 1 | admin1Name, data = data, na.action = na.omit)
summary(model.1, corr = FALSE)



```

```{r}

fixef(model.1)

```

```{r}

model.0 = lm(Mortality ~ Year  + Biomass_anomaly + Area, data = data, na.action = na.omit)
test=-2*logLik(model.0) +2*logLik(model.1)
mean(pchisq(test,df=c(0,1),lower.tail=F))
```

This shows that the random effect is necessary.

```{r}

model.2 = lme(Mortality ~ Year  + Biomass_anomaly + Area, random = list(admin1Name = pdDiag(~admin0Name)), data = data, na.action = na.omit)
summary(model.2, corr = FALSE)




```

```{r}

fixef(model.2)

```


```{r}

model.3 = lme(Mortality ~ Year  + Biomass_anomaly + Area, random = ~ 1 | admin0Name/admin1Name, data = data, na.action = na.omit)
summary(model.3, corr = FALSE)


```


```{r}

fixef(model.3)

```


```{r}

model.4 = update(model.2, .~.+Year:Biomass_anomaly+Area:Biomass_anomaly)
model.5 = update(model.3, .~.+Year:Biomass_anomaly+Area:Biomass_anomaly)


```

```{r}

anova(model.4,model.5)

```

```{r}
#Now model 4 versus model 1.
model.1b = update(model.1, .~.+Year:Biomass_anomaly+Area:Biomass_anomaly)
test=-2*logLik(model.1b) +2*logLik(model.4)
mean(pchisq(test,df=c(0,1),lower.tail=F))

```

```{r}
model.4 = update(model.4, method = 'ML')
model.2 = update(model.2, method = 'ML')
anova(model.2,model.4)

```

```{r}
#Model final?
model.2 = update(model.2, .~.-Biomass_anomaly, method = 'REML')
summary(model.2)
```


```{r}

fixef(model.2)

```

```{r}

plot(model.2 , Mortality ~ fitted(.) | admin0Name, abline = c(0,1))
plot(model.2 , Mortality ~ fitted(.) | admin1Name, abline = c(0,1))

```

```{r}

plot(model.2)

```
Not constant?

```{r}

model.2b=update(model.2,weights=varPower(form=~fitted(.)))

plot(model.2)

```


```{r}

qqnorm(resid(model.2b))
qqline(resid(model.2b))
shapiro.test(resid(model.2b))


```

Again we try to transform the response variable to achieve normality in the residuals:

```{r}


data$Mortality = sqrt_x(data$Mortality)$x.t

model.final = lme(Mortality ~ Year + Area, random = list(admin1Name = pdDiag(~admin0Name)), data = data, na.action = na.omit)


summary(model.final)
plot(model.final)
qqnorm(resid(model.final))
qqline(resid(model.final))
shapiro.test(resid(model.final))

```

```{r}

plot(model.final , Mortality ~ fitted(.) | admin0Name, abline = c(0,1))
plot(model.final , Mortality ~ fitted(.) | admin1Name, abline = c(0,1))

```


```{r}

plot(ranef(model.2))

```